Calculating life expectancy and exploring missingness

### 2016-10-18 Code/1_Explore_Data_v2.Rmd
added to the age decomp fx and fixed some coding errors

@@

```{r create_dataset_BlackWhite}
Black <- dat2 %>% filter(Race2 == "Black" & Age3 == 0) %>% 
                  select(State2, Year3, Sex2, Pop_across_age, life.expectancy.birth) %>%
                  rename(Pop_black = Pop_across_age,  le.birth.black = life.expectancy.birth)

@@

```{r calculate_age_decomp}

#create list of stratum.ids for each race-sex combination
for (rs in unique(levels(dat.clean$RaceSex))) {
  assign(paste0(gsub(" ", "", rs), ".stratum.ids"), 
      dat.clean %>% filter(dat.clean$RaceSex == rs) %>%
        select(StateYearRaceSex, StateYearSex, Sex2, State2, Year3) %>%
        group_by(StateYearRaceSex, StateYearSex) %>%
        filter(row_number() == 1)# %>%
      #  rename(paste0(gsub(" ", "", rs), ".ids") = StateYearRaceSex)
         )

@@

#merge the black-white lists
Female.ids <- merge(WhiteFemale.stratum.ids, BlackFemale.stratum.ids[, -c(3:5)], 
                    by = "StateYearSex", suffixes = c("White", "Black"))
head(Female.ids)

@@

paired.ids <- rbind(Male.ids, Female.ids)

@@

for(i in 1:dim(paired.ids)[1]) {
  white.index <- which(non.zero.stratums == paired.ids[i, 2])
  black.index <- which(non.zero.stratums == paired.ids[i, 3])

>>>>

rm(les)
```

@@

```{r create_dataset_BlackWhite}
#this dataset contains the black-white data within the same row (wide format) for each year-state-sex strata

Black <- dat2 %>% filter(Race2 == "Black" & Age3 == 0) %>% 
                  select(State2, Year3, Sex2, Pop_across_age, life.expectancy.birth) %>%
                  rename(Pop_black = Pop_across_age,  le.birth.black = life.expectancy.birth)

@@

```{r create_paired_ids_dataset}
#paired.ids will contain the index names that are sent to the age decomposition function in the next code chunk

#first, create list of stratum.ids for each race-sex combination
for (rs in unique(levels(dat.clean$RaceSex))) {
  assign(paste0(gsub(" ", "", rs), ".stratum.ids"), 
         dat.clean %>% filter(dat.clean$RaceSex == rs) %>%
         select(StateYearRaceSex, StateYearSex, Sex2, State2, Year3) %>%
         group_by(StateYearRaceSex, StateYearSex) %>%
         filter(row_number() == 1)
         )

@@

#then, merge the black-white lists
Female.ids <- merge(WhiteFemale.stratum.ids, BlackFemale.stratum.ids[, -c(3:5)], 
                    by = "StateYearSex", suffixes = c("White", "Black"))
head(Female.ids)

@@

paired.ids <- rbind(Male.ids, Female.ids)
```

```{r calculate_age_decomp}
#runs in about a minute

@@

for(i in 1:dim(paired.ids)[1]) {
  white.index <- which(non.zero.stratums == paired.ids[i, "StateYearRaceSexWhite"])
  black.index <- which(non.zero.stratums == paired.ids[i, "StateYearRaceSexBlack"])
